---
title: "Lab 1: R-Basics"
output:
  html_document:
    css: projectlab.css
    highlight: pygments
    theme: yeti
    toc: true
    toc_float: true
    toc_depth: 4
editor_options: 
  markdown: 
    wrap: sentence
---

```{r, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, 
                      fig.show = "hide", 
                      message = FALSE, 
                      warning = FALSE)

library(openintro)
library(tidyverse)
library(broom)
library(AER)
library(readxl)
library(kableExtra)
LabRubric <- read_excel("Table_LabRubric.xlsx")
```

\

# Overview

\
**Welcome to your GEOG-364 labs!**

The aim of these labs is to help you get comfortable programming in R and to practice techniques introduced in lectures and homework.

\

::: {#boxedtext}
**Questions**

Can you use Chat-GPT/AI?
What’s the late policy?
Where can you get help?\
Click here for the answers.
:::

\

------------------------------------------------------------------------

\

# Lab 1

**Aim of the lab**

This lab builds on Homework 1 from DataCamp.
Your goal is to become familiar with R and R Markdown, and to learn about some of their core features, including:

-   Packages and libraries
-   R Markdown documents
-   Basic code commands
-   Loading data
-   Summary statistics for different types of data
-   How labs are structured and what you need to submit

In Lab 2, we will explore similar ideas but with spatial data.

\

**Deadline & Late Penalty**

-   Tuesday lab: to add
-   Thursday lab: to add
-   Friday lab: to add

\

**What to submit**

You will submit BOTH your Rmd and HTML files.
This will make sense by the end!

\

**How long should the lab take?**

To add.

\

**Getting help**

-   Kiely (and often Dr G) will be present during your lab sessions. This is the easiest way to get help.
-   There is a Piazza discussion board. Kiely will explain more in this week’s lab and will check it at least once a day.
-   Dr G has weekly office hours and genuinely enjoys helping with R, even if you feel stuck or overwhelmed.
-   If you have posted on Piazza and received no response after 24 hours, you may send a Canvas message to Kiely or Dr G (or if you are completely lost).\

------------------------------------------------------------------------

# Getting Started

ksadnasdj

adhjladsnjkad

<div>

**"Some boxed text"**

Lalalalal lalala alalala

</div>

------------------------------------------------------------------------

<br>

# Basic data analysis

asdjlasdld asdjlasdld

asdjlasdld asdjlasdld

\

asdjlasdld asdjlasdld

<br>

asdjlasdld asdjlasdld

<br>

------------------------------------------------------------------------

<br>

# Logistic regression vs mother's weight

adadmklasdm

<br>

------------------------------------------------------------------------

# Show me something new!

adasdsad

<br>

------------------------------------------------------------------------

# What to submit?

asdjnasda

<br>

# Grading rubric

## Overall

Overall, here is what your lab should correspond to:

```{r, echo=FALSE}
rubric <- readxl::read_excel("Table_LabRubric.xlsx")
knitr::kable(rubric) %>%   
  kable_classic_2() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))

```

And..
finished!
