---
title: "Lab 4: TBD"
output:
  html_document:
    css: projectlab.css
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
editor_options: 
  markdown: 
    wrap: sentence
---

```{r, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, 
                      fig.show = "hide", 
                      message = FALSE, 
                      warning = FALSE)

library(openintro)
library(tidyverse)
library(broom)
library(AER)
library(readxl)
library(kableExtra)
LabRubric <- read_excel("Table_LabRubric.xlsx")
```

# Overview

Welcome to your labs. They are designed to help you move from the basics in R, to do some complex spatial analysis. 

**Style**

Your style should be clear, understandable, explaining any jargon as simply as you can. You are also welcome to include in hand-drawn sketches, or CITED diagrams from the internet.
<br><br>

**What to submit**

You will submit BOTH your RmD and html files. This will make sense by the end!

<br>

------------------------------------------------------------------------

# Getting Started

ksadnasdj

adhjladsnjkad

::: {#boxedtext}
**"Some boxed text"**

Lalalalal 
lalala
alalala
:::

------------------------------------------------------------------------

<br>

# Basic data analysis


asdjlasdld

<br>

------------------------------------------------------------------------

<br>


# Logistic regression vs mother's weight

adadmklasdm

<br>




------------------------------------------------------------------------

# Show me something new!

adasdsad

<br>


------------------------------------------------------------------------

# What to submit?

asdjnasda

<br>



# Grading rubric



## Overall

Overall, here is what your lab should correspond to:

```{r, echo=FALSE}
rubric <- readxl::read_excel("Table_LabRubric.xlsx")
knitr::kable(rubric) %>%   
  kable_classic_2() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))

```


And.. finished!







