---
title: "Lab 1: R-Basics"
output:
  html_document:
    css: projectlab.css
    highlight: pygments
    theme: yeti
    toc: true
    toc_float: true
    toc_depth: 4
editor_options: 
  markdown: 
    wrap: sentence
---

```{r, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, 
                      fig.show = "hide", 
                      message = FALSE, 
                      warning = FALSE)

library(openintro)
library(tidyverse)
library(broom)
library(AER)
library(readxl)
library(kableExtra)
LabRubric <- read_excel("Table_LabRubric.xlsx")
```

<br>

# Overview

**Welcome to your GEOG-364 labs!**  

Overall, the goals of these labs are to get you used to programming in R and to try some of the techniques that we talk about in the lectures and homeworks.

\

::: {#boxedtext}
**Questions**

Can you use Chat-GPT/AI?  What's the late policy?  Where can I get help?
Click here to find out the answers.

:::


<br>

------------------------------------------------------------------------

<br>

# Lab 1

**Aim of the lab**

This lab builds on your Homework 1 from Data Camp. Your aim is to start getting comfortable in R and R-Markdown, and to learn about some of its core features including:

 - Packages/libraries
 - R-Markdown & making websites
 - Basic code commands
 - Loading data
 - Summary statistics for different types of data
 - Understanding lab structures and what to submit
 
In lab 2, we will then move onto similar things but for spatial data. 
 
<br>

**What to submit**

You will submit BOTH your RmD and html files. This will make sense by the end!

<br>

------------------------------------------------------------------------

# Getting Started

ksadnasdj

adhjladsnjkad

::: {#boxedtext}
**"Some boxed text"**

Lalalalal 
lalala
alalala
:::

------------------------------------------------------------------------

<br>

# Basic data analysis

asdjlasdld
asdjlasdld

asdjlasdld
asdjlasdld

\

asdjlasdld
asdjlasdld

<br>

asdjlasdld
asdjlasdld

<br>

------------------------------------------------------------------------

<br>


# Logistic regression vs mother's weight

adadmklasdm

<br>




------------------------------------------------------------------------

# Show me something new!

adasdsad

<br>


------------------------------------------------------------------------

# What to submit?

asdjnasda

<br>



# Grading rubric



## Overall

Overall, here is what your lab should correspond to:

```{r, echo=FALSE}
rubric <- readxl::read_excel("Table_LabRubric.xlsx")
knitr::kable(rubric) %>%   
  kable_classic_2() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))

```


And.. finished!







