---
title: "Welcome to GEOG-364"
subtitle: <h4 style="font-style:normal">GEOG-364 - Spatial Analysis</h4>
author: Penn State, Dept Geography
output: 
  html_document:
    toc: false
    theme: flatly
    toc_float:
      collapsed: no 
      smooth_scroll: yes
---


```{r setup, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
# Chunk options
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE)

## Load Libraries
library(tidyverse)
library(dplyr)
library(plotly)
library(kableExtra)

## Load Data
rubric <- readxl::read_excel("./index_data/Table_LabRubric.xlsx")
```


<br>

# Welcome to GEOG-364! {#Home .unnumbered}

<br>

Dr Greatrex has written this lab book to provide all the code and instructions you need for spatial analysis. You will be conducting analysis using R and writing up your reports using R-Markdown.

[Unless I tell you otherwise, you can find all the code you need in this lab book]{.underline}

Don't panic if you are new to this! We will start from scratch and you are expected to have zero programming experience.


LINKS NEED UPDATING

<br>

## Useful Links {#Home_UsefulLinks .unnumbered}

-   [The course canvas page:]{.underline} <https://psu.instructure.com/courses/2312605>

-   [Have a question?]{.underline} (I'm late! Can I work with friends? What about ChatGPT? etc), see [Lab FAQs](#FAQ "Lab FAQs")

-   [Brand new?]{.underline} Start here: [What is R](#WhatIsR), then go here if you want to use it on your own computer [Installing for the first time](#WhatIsR) or here if you want to use it online.

-   [Used R before?]{.underline} Please update your version. Instructions and the reason why here: [Updating R and R-Studio](#Setup_UpdateDesktop), Click here to read here for more on a tool you might not have used, [R-Markdown](#WhatisMarkdown).

<br>

## How to use this website {#Home_howtouse .unnumbered}

This website has been written using R-bookdown and uploaded to github. The format has been tweaked using the excellent bookdown example of [Mastering Shiny](https://mastering-shiny.org/).

Expand here to see how to navigate:

<br>


```{r, eval=FALSE, include=FALSE}
# automatically create a bib database for R packages
#knitr::write_bib(c(
#  .packages(), 'bookdown', 'knitr', 'rmarkdown'
#), 'packages.bib')
```
