---
title: "GEOG-364 Lab book"
subtitle: <h4 style="font-style:normal">GEOG-364 - Spatial Analysis</h4>
author: Penn State, Dept Geography
output: 
  html_document:
    toc: false
    theme: flatly
    toc_float:
      collapsed: no 
      smooth_scroll: yes
---


```{r setup, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
# Chunk options
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE)

## Load Libraries
library(tidyverse)
library(dplyr)
library(plotly)
library(kableExtra)

## Load Data
rubric <- readxl::read_excel("./index_data/Table_LabRubric.xlsx")
```


<br>


### Welcome!

Dr Greatrex has written this lab book to provide all the code and instructions you need for spatial analysis. You will be conducting analysis using R and writing up your reports using R-Markdown.

Don't panic if you are new to this! We will start from scratch and you are expected to have zero programming experience. [Unless I tell you otherwise, you can find all the code you need in this lab book and my tutorials]{.underline}

-   [Lab 1:](#Lab1 "Lab 1")


## Useful Links {#Home_UsefulLinks .unnumbered}

-   [**The course canvas page:**](https://psu.instructure.com/courses/2396422)

-   [**Frequently asked questions** (I'm late! Can I work with friends? What about ChatGPT?)](#FAQ "Lab FAQs")

-   [**All the tutorials**](https://psu-spatial.github.io/Stat462-2025/CH1_BASICS.html#ch1)

<br>



```{r, eval=FALSE, include=FALSE}
# automatically create a bib database for R packages
#knitr::write_bib(c(
#  .packages(), 'bookdown', 'knitr', 'rmarkdown'
#), 'packages.bib')
```
